Realizing Temporal Graphs

### 2023-01-19 RealizingTemporalGraphs.tex
Changes Wednesday 18th of Jan

    \begin{observation}\label{obs:travel-delays-both-directions}
    Let $(G, \lambda, \Delta)$ be a temporal graph satisfying conditions of problem \deltaExact.
    For any two incident edges $e_1 = uv$ and $e_2 = vz$ on vertices $u,v,z \in V$, with $e_1 \cap e_2 = v$, we have $\tau_v^{zu} = \Delta - \tau_v^{uz} \pmod \Delta$.
    \end{observation}

@@

Note that we can easily solve the \deltaExact\ by guessing the structure of temporal paths among all pairs of vertices.
But, this would create $f(n,m)$ guesses, and the running time of the algorithm would be too big.
Therefore, we approach the problem using some special properties, which allow us to create $f(k)$ different guesses.
We solve each guess by creating an ILP instance with $f(k)$ variables and $O(n^c)$ constraints, for some $c \in \mathbb{N}$.
The ILP instance is a \textsc{Yes} instance if and only if $D$ is a \textsc{Yes} instance of \deltaExact.
Moreover, from a positive solution of the ILP instance we can construct the desired labeling $\lambda$.
Using results from \cite{Lenstra1983Integer} we know that an ILP instance can be solved in FPT time with respect to the number of variables.
Therefore we can solve each of our ILP instances in $f(k)$ time.

We now introduce the algorithm, that determines the labeling of $G$, satisfying $D$, and runs in $f(k) \cdot \text{poly}(n)$ time.
Our algorithm performs the following steps.
\begin{enumerate}
    \item Preprocessing on graph $G$. Determine the set $E$ of feedback edges, the set $U$ of vertices of interest and the set $U^*$ of neighbors of vertices of interest, where $|E|, |U|, |U^*| \in O(k)$.
    \item Guess what are the fastest temporal paths among vertices in $U$ and 
    what are the fastest temporal paths among the vertices in $U^*$, that are of some specific form.
    \begin{enumerate}
        \item For each guess create an ILP instance $I$ with $f(k)$ variables.
        \item Add constraints for all pairs of vertices.
    \end{enumerate}
    \item Solve the ILP instance $I$ and check if a solution exists. In the positive case create its corresponding labeling $\lambda_I$ and check if it satisfies $D$.
    All in total we create $f(k)$ partial ILP instances, each with $f(k)$ variables and $f(n)$ constraints.
\end{enumerate}

\subsection{Step 1 of the FPT algorithm}
From the underlying graph $G$ of $D$ we first create a graph $G'$ by
iteratively removing vertices of degree one from $G$.
Then we determine a minimum feedback edge set $F$ of $G'$ by finding a spanning tree $T$ of $G'$ and set $F = E(G) \setminus E(T)$. 
Note that $F$ is also a minimum feedback edge set of the whole $G$.
Lastly, we determine sets $U$, of \emph{vertices of interest}, and $U^*$ of the neighbors of vertices of interest, in the following way.
Let $T$ be a spanning tree of $G'$, with $F$ being the corresponding feedback edge set of $G'$.
Let $V_1 \subset V(G')$ be the set of leaves in the spanning tree $T$, $V_2 \subset V(G')$ be the set of vertices of degree two in $T$, that are incident to at least one edge in $F$, 
and let $V_3 \subset V(G')$ be the set of vertices of degree at least $3$ in $T$. 
Then $|V_1| + |V_2| \leq 2k$, since every leaf in $T$ and every vertex in $V_2$ is incident to at least one edge in $F$,
and $|V_3| \leq |V_1|$ by the properties of trees.
We denote with $U = V_1 \cup V_2 \cup V_3$ the set of vertices of interest. It follows that $|U| \leq 4k$.
We set $U^*$ to be the set of vertices in $V(G') \setminus U$ that are neighbors of vertices in $U$, \ie $U^* = \{v \in V(G') \setminus U | u \in U, v \in N(u)\}$.
Again, using the tree structure, we get that for any $u \in U$ its neighborhood is of size $|N(u)| \in O(k)$, therefore $U^* \in O(k^2)$.

@@

\subsection{Step 2 of the FPT algorithm}
Let us denote with $Z$ the set of vertices removed from $G$ at the beginning of the algorithm, \ie $Z = V(G) \setminus V(G')$.
By the construction it follows that $Z$ consists of disjoint trees. 
Note that the fastest temporal paths among vertices of $G'$ are independent of vertices in $Z$.
Let us remember the following basic property of the problem \deltaExact.
A labeling $\lambda$ of $G$ satisfies $D$ if every fastest temporal path from vertex $v_i$ to $v_j$ equals $D_{v_i, v_j}$,
and none other temporal path from $v_i$ to $v_j$ is faster.
In order to find a labeling that satisfies this property we split our analysis in nine cases,
where the first vertex is in one of the sets $U,V(G') \setminus U, Z$,
and similarly the last vertex is in one of the sets $U,V(G') \setminus U, Z$.
In each of these cases we guess the underlying path $P$ of the fastest temporal path from the vertex $v_i$ to $v_j$, 
which results in one equality constraint for the labels on the path $P$. 
For all other temporal paths from $v_i$ to $v_j$ we know that they cannot be faster, so we introduce inequality constraints for them.
This results in producing $f(n,m)$ constraints. We have to guarantee only, that the number of guesses, of faster temporal paths, is in $f(k)$ and not in $f(n)$.

We start the algorithm by guessing the structure of 
the fastest temporal paths (we guess the underlying paths) among all pairs of vertices of $U$. 
For a pair $u,v$ of vertices in $U$, there are $k!$ possible paths in $G'$ between them. 
Therefore, we have to try all $O(k^k)$ possible paths, where at least one of them will be a fastest temporal path from $u$ to $v$, respecting the values from $D$.
If this does not happen we conclude that $D$ is a \textsc{No} instance of \deltaExact.
Repeating this procedure for all pairs of vertices $u,v \in U$ we get $O((k^k)^{k^2})=O(k^{k^3})$ different variations of the fastest temporal paths among all vertices in $U$.

We now create one partial ILP instance for each guess of fastest temporal paths among the vertices from $U$,
which we later on extend to $f(k)$ final ILP instances, after guessing some extra fastest temporal paths.
From now on, for every pair of vertices $u,v \in U$ we assume that we know exactly what is the underlying path of a fastest temporal path from $u$ to $v$.

For an easier understanding and the analysis of the algorithm we give the following definition.
\begin{definition}
    Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
    A path $P = (u=v_1,v_2, \dots, v_p = v)$ in $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for $i \in \{ 2, 3, \dots, p-1\}$,
    is called a \emph{segment}. Denote it with $S_{u,v}$.
\end{definition}

We describe now how to build one partial ILP instance from a specific list of fastest temporal paths.
While building the ILP instances we exploit the following property greatly, that is why we state it as a separate observation.

\begin{observation}\label{obs:FirstLabelAndDuration}
    Let $P$ be the underlying path of a fastest temporal path from $u$ to $v$, where $e_1, e_p \in P$ are its first and last edge, respectively.
    Then, knowing the label $\lambda (e_1)$ of the first edge and the duration $d(P)$ of $P$, we can uniquely determine the label $\lambda (e_p)$ of the last edge of $P$.
    Symmetrically, knowing $\lambda(e_p)$ and $d(P)$, we can uniquely determine $\lambda(e_1)$.
\end{observation}
The correctness of the above statement follows directly from \cref{def:temporalPath+Duration}. Since the duration of $P$ is calculated as the difference of labels of last and first edge plus $1$,
where the label of last edge is considered with some delta periods,
\ie $d(P) = p_i \Delta + \lambda(e_p) - \lambda (e_1) + 1$, for some $p_i \geq 0$.
Therefore $d(P) \pmod \Delta \equiv  (\lambda(e_p) - \lambda (e_1) + 1) \pmod \Delta$.

@@

\begin{claim}\label{claim:FPT-uv-Labelalledges}
    Let $u, v \in U$ be two arbitrary vertices of interest and suppose that $P = (u=v_1,v_2, \dots, v_p = v)$, where $p \geq 2$, 
    is a $u,v$ path in $G'$, which is also the underlying path of a fastest temporal path from $u$ to $v$.
    Suppose also that no other vertex from $P$ is a vertex of interest, \ie $P \setminus \{u,v\} \cap U = \emptyset$.
    We can determine the labeling $\lambda$ of every edge in $P$, up to the constant $c_{uv} = \lambda(uv_2)$,
    such that the labeling $\lambda$ respects $D'$.
\end{claim}

\begin{proof}
In this case we get that all other vertices in $P$ are reached from $u$ using a (sub)path of $P$.
To prove this suppose for the contradiction that there is a vertex $v_i \in P \setminus \{u,v\}$, that is reached from $v$ on a path different than $P_i = (u, v_2, v_3, \dots, v_i)$.
Since the only vertices of interest of $P$ are $u$ and $v$, it follows that all other vertices on $P$ are of degree $2$. 
Then the only way to reach $v_i$ from $u$, that differs from $P$, would be to go from $u$ to $v$ using a different path $P_2$,
and then go from $v$ to $v_{p-1}, v_{p-2}, \dots, v_i$.
But since $P$ is the fastest temporal path from $u$ to $v$, we get that $d(P_2) \geq d(P)$ and $d(P_2 + (v,v_{p-1}, \dots, v_i) > d(P) > d(P_i)$.

@@

We use the above results in the following way.
Let $u, v \in U$ be two vertices of interest. 
We know the underlying path $P_{u,v}$ of a fastest temporal path from $u$ to $v$. 
Suppose that $P_{u,v}$ consist of $\ell$ segments $S_{u_i,v_i}$.
Using \cref{claim:FPT-uv-LabelAlmostalledges} we determine the labels of all but $\ell -1$ edges on $P_{u,v}$, with respect to the value of the first edge.
This results in creating one constraint, by setting the duration of the temporal path along $P_{u,v}$ to be equal to the value $D_{u,v}$.
There are $O(k^k)$ different paths $P^i_{u,v}$ from $u$ to $v$ in $G$.
We now have to iterate through all of them and add the constraint that the duration of $(P^i_{u,v}, \lambda)$ is greater or equal to $D_{u,v}$.
We repeat the above procedure for all pairs of vertices in $U$.
This results in producing $O(k^{k^3})$ variables and constraints to our partial ILP.

\begin{comment}
We create a variable $c_{u,v}^{e_1} \in [\Delta]$ that corresponds to the label of the first edge of $P_{u,v}$ in $(G', \lambda')$ and potentially at most $\ell_{u,v}$ variables $c_{u,v}^{e_w} \in [\Delta]$ corresponding to 
labels of edges $e_w \in P_{u,v}$ that are incident to a vertex of interest $w$, for which the underlying path $P_{u,w}$ of a fastest temporal path from $u$ to $w$ intersects $P_{u,v}$ only in vertices $u$ and $w$.
Labels on all other edges of $P_{u,v}$ are determined exactly, in relation to the constant $c_{u,v}^{e_1}$.
Repeating this procedure for all $u,v \in U$ creates at most $|U|^3 = O(k^3)$ new variables.
\end{comment}
\todo[inline]{Write examples for variables and constrains this part produces.}

@@

We now continue with guessing fastest temporal paths among all vertices in $U^*$, 
which by similar arguing as for vertices in $U$, gives us $O(k^{k^6})$ guesses.
Besides that we also guess fastest temporal paths of the following form.
Let $S_{u,v} = (u=v_1,v_2, \dots, v_p = v)$ and $S_{w,z} = (w=z_1,z_2, \dots, z_r = z)$ be two segments,
first one between vertices $u,v \in U$, second one between vertices $w, z \in U$, 
where the first one includes $v_2, v_{p-1} \in U^*$, 
and the second one includes $z_2, z_{p-1} \in U^*$.
We want to guess the underlying paths of the fastest temporal paths from $v_2$ to $z_2$,
where the first path goes 
from $v_2$, through $u$, through $w$ and then to $z_2$,
and the second one goes
from $v_2$, through $u$, through $z$ and then to $z_2$.
And two similar paths from $v_2$ to $z_{p-1}$.
We repeat this procedure for all pairs of segments and all vertices from $U^*$ in those segments.
Since there are $O(k^3)$ segments in $G'$,
there are $O(k^6)$ possible paths of this form.
Lastly,
for a segment $S_{u,v} = (u=v_1,v_2, \dots, v_p = v)$
and a vertex of interest $w \in U$,
we guess the underlying paths of the fastest temporal paths from $w$ to $v_2$, through $u$ and through $v$.
Besides that we guess the underlying paths of the fastest temporal paths from $v_2$ to $w$ through $u$ and through $v$.
For a fixed segment $S_{u,v}$ and a fixed vertex of interest $w$ we have $O(4 k^k)$ different possible such paths, therefore we make $O(k^6)$ guesses for these paths.
All together we made $O(k^6)$ new guesses for the structure of all of the above mentioned fastest temporal paths.
Each of the previous partial ILP instances is therefore extended into $O(k^6)$ new ILP instances.

>>>>

    From the above definition and the definition of the duration of the temporal path $P$ we get the following two observations.
    \begin{observation}\label{obs:durationPwithWaitingTimes}
        Let $P = (v_0, v_1, \dots, v_p)$ be the temporal path from $v_0$ to $v_p$..
        Then $d(P) = \sum_{i = 1}^{p-1} \tau_{v_i}^{v_{i-1}v_i} + 1 $.
    \end{observation}
    \begin{proof}
        For the simplicity of the proof denote $t_i = \lambda(v_{i-1}v_i)$, and suppose that $t_i \leq t_{i+1}$, for all $i \in \{1,2,3,\dots,p\}$.
        Then
        \begin{align*}
        \sum_{i = 1}^{p-1} \tau_{v_i}^{v_{i-1}v_i} + 1  
        &= \sum _{i = 1}^{p-1} (t_{i+1} - t_i) + 1 \\
        & = (t_2 - t_1) + (t_3  t_2) + \cdots + (t_p - t_{p-1}) + 1  \\
        & = t_{p-1} - t_1 + 1\\
        & = d(P)
        \end{align*}
        Now in the case when $t_i > t_{i+1}$ we get that $\tau_{v_i}^{v_{i-1}v_{i+1}} = \Delta + t_{i+1} - t_i$.
        At the end this still results in the correct duration as the last time we traverse the path $P$ is not exactly $t_p$ but $k \lambda + t_p$, for some $k$.
    \end{proof}
    We also get the following.
    \begin{observation}\label{obs:travel-delays-both-directions}
    Let $(G, \lambda, \Delta)$ be a temporal graph satisfying conditions of problem \deltaExact.
    For any two incident edges $e_1 = uv$ and $e_2 = vz$ on vertices $u,v,z \in V$, with $e_1 \cap e_2 = v$, we have $\tau_v^{zu} = \Delta - \tau_v^{uz} \pmod \Delta$.
    \end{observation}

@@

Note that we can easily solve the \deltaExact\ by guessing the structure of temporal paths among all pairs of vertices.
But, this would create $f(n,m)$ guesses (where $f(n,m)$ is an exponential function), and the running time of the algorithm would be too big.
Therefore, we approach the problem using some special properties, which allow us to create $f(k)$ different guesses.
We solve each guess by creating an ILP instance with $O(k^3)$ variables and $O(n^c)$ constraints, for some $c \in \mathbb{N}$.
The ILP instance is a \textsc{Yes} instance if and only if $D$ is a \textsc{Yes} instance of \deltaExact.
Moreover, from a positive solution of the ILP instance we can construct the desired labeling $\lambda$.
Using results from \cite{Lenstra1983Integer} we know that an ILP instance can be solved in FPT time with respect to the number of variables.
Therefore we can solve each of our ILP instances in FPT time with respect to $k$.

We now introduce the algorithm, that determines the labeling of $G$, satisfying $D$, and runs in $f(k) \cdot \text{poly}(n)$ time.
Our algorithm performs the following steps.
\begin{enumerate}
    \item Preprocessing on graph $G$. Determine the set $E$ of feedback edges, the set $U$ of vertices of interest and the set $U^*$ of neighbors of vertices of interest, where $|E|, |U|, |U^*| \in O(k)$.
    \item Guess what are the fastest temporal paths among vertices in $U$ and 
    what are the fastest temporal paths among the vertices in $U^*$, that are of some specific form.
    \begin{enumerate}
        \item For each guess create an ILP instance $I$ with $f(k)$ variables.
        \item Add constraints for all pairs of vertices.
    \end{enumerate}
    \item Solve the ILP instance $I$ and check if a solution exists. In the positive case create its corresponding labeling $\lambda_I$ and check if it satisfies $D$.
    All in total we create $f(k)$ partial ILP instances, each with $f(k)$ variables and $O(n^c)$ constraints.
\end{enumerate}

\subsection{Preprocessing of the input}
From the underlying graph $G$ of $D$ we first create a graph $G'$ by
iteratively removing vertices of degree one from $G$.
Then we determine a minimum feedback edge set $F$ of $G'$ by finding a spanning tree $T$ of $G'$ and set $F = E(G) \setminus E(T)$. 
Note that $F$ is also a minimum feedback edge set of the whole $G$.
Lastly, we determine sets $U$, of \emph{vertices of interest}, and $U^*$ of the neighbors of vertices of interest, in the following way.
Let $T$ be a spanning tree of $G'$, with $F$ being the corresponding feedback edge set of $G'$.
Let $V_1 \subset V(G')$ be the set of leaves in the spanning tree $T$, $V_2 \subset V(G')$ be the set of vertices of degree two in $T$, that are incident to at least one edge in $F$, 
and let $V_3 \subset V(G')$ be the set of vertices of degree at least $3$ in $T$. 
Then $|V_1| + |V_2| \leq 2k$, since every leaf in $T$ and every vertex in $V_2$ is incident to at least one edge in $F$,
and $|V_3| \leq |V_1|$ by the properties of trees.
We denote with $U = V_1 \cup V_2 \cup V_3$ the set of vertices of interest. It follows that $|U| \leq 4k$.
We set $U^*$ to be the set of vertices in $V(G') \setminus U$ that are neighbors of vertices in $U$, \ie $U^* = \{v \in V(G') \setminus U | u \in U, v \in N(u)\}$.
Again, using the tree structure, we get that for any $u \in U$ its neighborhood is of size $|N(u)| \in O(k)$, therefore $U^* \in O(k^2)$.

@@

Let us denote with $Z$ the set of vertices removed from $G$ at the beginning of the algorithm, \ie $Z = V(G) \setminus V(G')$.
By the construction it follows that $Z$ consists of disjoint trees. 
Note that the fastest temporal paths among vertices of $G'$ are independent of vertices in $Z$.
Let us remember the following basic property of the problem \deltaExact.
A labeling $\lambda$ of $G$ satisfies $D$ if every fastest temporal path from vertex $v_i$ to $v_j$ equals $D_{v_i, v_j}$,
and none other temporal path from $v_i$ to $v_j$ is faster.
In order to find a labeling that satisfies this property we split our analysis in nine cases,
where the first vertex is in one of the sets $U,V(G') \setminus U, Z$,
and similarly the last vertex is in one of the sets $U,V(G') \setminus U, Z$.
In each of these cases we guess the underlying path $P$ of the fastest temporal path from the vertex $v_i$ to $v_j$, 
which results in one equality constraint for the labels on the path $P$. 
For all other temporal paths from $v_i$ to $v_j$ we know that they cannot be faster, so we introduce inequality constraints for them.
This results in producing $poly(n,m)$ constraints. We have to guarantee only, that the number of guesses, of faster temporal paths, is exponential in $k$ and not in $n$.

For an easier understanding and the analysis of the algorithm we give the following definition.
\begin{definition}
    Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
    A path $P = (u=v_1,v_2, \dots, v_p = v)$ in $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for $i \in \{ 2, 3, \dots, p-1\}$,
    is called a \emph{segment}. Denote it with $S_{u,v}$.
\end{definition}
From the definition it follows, that there are exactly $k^2$ segments in $G'$.
Observe also that a temporal path in $G'$ between two vertices of interest is either a segment, or consists of a sequence of some segments.
%For a pair $u,v$ of vertices in $U$, there are $O(k!)=O(k^k)$ possible paths in $G'$ between them.

\subsection{Creating ILP instances and determining variables}
For every edge $e$ in $G'$ that is incident to a vertex of interest we create a variable $x_e$ that can have values from $ \{1, 2, \dots, \Delta\}$.
Each vertex of interest is incident to at most $k$ such edges, therefore we create $O(k^3)$ variables.
We start the algorithm by guessing necessary structures.
Note, whenever we say that we guess the fastest temporal path between two vertices, we mean that we guess the underlying path of the fastest temporal path among those two vertices.
In the case when we want to guess the fastest path from $u$ to $v$, that passes through the vertex $x$ we write it as a fastest path of form $u \rightarrow x \rightarrow v$.
We guess the following structures.
\begin{enumerate}
    \item The fastest temporal paths among all pairs of vertices of $U$.
    For a pair $u,v$ of vertices in $U$, there are $k!$ possible paths in $G'$ between them. 
    Therefore, we have to try all $O(k^k)$ possible paths, where at least one of them will be a fastest temporal path from $u$ to $v$, respecting the values from $D$.
    %If this does not happen we conclude that $D$ is a \textsc{No} instance of \deltaExact.
    Repeating this procedure for all pairs of vertices $u,v \in U$ we get $O((k^k)^{k^2})=O(k^{k^3})$ different variations of the fastest temporal paths among all vertices in $U$.
    \item The fastest temporal paths among all vertices in $U^*$, 
    which by similar arguing as for vertices in $U$, gives us $O(k^{k^6})$ guesses.
    \item The fastest temporal paths from vertices of $U$ to vertices in $U^*$,
    and vice versa, the fastest temporal paths from vertices in $U^*$ to vertices in $U$.
    This gives us $O(k^{k^5})$ guesses.
    \item Let $S_{u,v} = (u=v_1,v_2, \dots, v_p = v)$ and $S_{w,z} = (w=z_1,z_2, \dots, z_r = z)$ be two segments,
    first one between vertices $u,v \in U$, second one between vertices $w, z \in U$, 
    where the first one includes $v_2, v_{p-1} \in U^*$, 
    and the second one includes $z_2, z_{p-1} \in U^*$.
    We want to guess the following fastest temporal paths
    $v_2 \rightarrow u \rightarrow w \rightarrow z_2$,
    $v_2 \rightarrow u \rightarrow z \rightarrow z_2$,
    $v_2 \rightarrow u \rightarrow w \rightarrow z_{r-1}$ and
    $v_2 \rightarrow u \rightarrow z \rightarrow z_{r-1}$.
    We repeat this procedure for all pairs of segments and all vertices from $U^*$ in those segments.
    Since there are $O(k^3)$ segments in $G'$,
    there are $O(k^6)$ possible paths of this form.
    \item For a segment $S_{u,v} = (u=v_1,v_2, \dots, v_p = v)$
    and a vertex of interest $w \in U$,
    we guess the following fastest temporal paths
    $w \rightarrow u \rightarrow v_2$, $w \rightarrow v \rightarrow v_2$,
    and
    $v_2 \rightarrow u \rightarrow w$, $v_2 \rightarrow v \rightarrow w$.
    For a fixed segment $S_{u,v}$ and a fixed vertex of interest $w$ we have $O(4 k^k)$ different possible such paths, therefore we make $O(k^6)$ guesses for these paths.
    All together we made $O(k^6)$ new guesses for the structure of all of the above mentioned fastest temporal paths.
%Each of the previous partial ILP instances is therefore extended into $O(k^6)$ new ILP instances.
    \item 
    Let $S_{u,v} = (u=v_1,v_2, \dots, v_p = v)$ be a line segment in $G'$, and let us
    fix a vertex $v_i \in S_{u,v} \setminus \{u,v\}$.
    Let 
    $S_{w,z} = (w=z_1,z_2, \dots, z_r = z)$ be another segments in $G'$.
    Note that all inner vertices of $S_{w,z}$ (\ie vertices in $S_{w,z} \setminus \{w,z\}$)
    can be reached from $v_i$ by some path that has to go through $w$ oz $z$.
    In fact, this is true for all paths starting in any vertex not in $S_{w,z}$.
    Therefore, the values of the duration of the fastest path from $v_i$ to vertices on 
    the segment
    $S_{w,z}$ have to increase up to the certain point, when traversing the segment coming through $w$ and coming through $z$.
    Namely,
    let $z_j \in S_{w,z} \setminus \{w,z\}$ be the vertex with the biggest value  $D_{v_i,z_j}$,
    we call this vertex a \emph{split vertex of $v_i$ in the segment $S_{wz}$}.
    Then $D_{v_i,z_2} < D_{v_i,z_3} < \dots < D_{v_i,z_j}$ and 
    $D_{v_i,z_{p-1}} < D_{v_i,z_{p-2}} < \dots < D_{v_i,z_j}$.
    From this it follows that the fastest temporal paths from $v_i$ to $z_2, z_3, \dots, z_{j-1}$ go through $w$,
    and 
    the fastest temporal paths from $v_i$ to $z_{p-1}, z_{p-2}, \dots, z_{j+1}$ go  through $z$.
    We now have to guess which vertex $w$ or $z$ is on a fastest temporal path from $v_i$ to $z_j$.
    Similarly,
    all fastest temporal paths starting at $v_i$ have to go either through $u$ or   through $v$,
    which also gives us two extra guesses for the fastest temporal path from $v_i$ to $z_j$.
    Therefore, all together we have $4$ possibilities on how the fastest temporal path from $v_i$ to $z_j$ starts and ends.
    For a fixed $S_{u,v}, v_i$ and $S_{w,z}$ we find the vertex $z_j$ in    polynomial time, 
    or determine that this kind of vertex does not exist.
    We then make four guesses where we determine how the fastest temporal path from $v_i$ to $z_j$ passes through vertices $u,v$ and $w,z$.
    We repeat this procedure for all pairs of segments,
    which results in producing $O(k^6)$ new guesses.
    \item At the end we also guess the permutation of all $O(k^3)$ variables, \ie for any two variables $x_e$ and $x_f$, we know if $x_e < x_f$ or $x_e = x_f$, or $x_e > x_f$.
    This results in $O(k^k)$ guesses.
\end{enumerate}
We create all of the guesses iteratively.
We start with the first one, that results in $O(k^{k^3})$ different possibilities, then we split each one of these guesses into $O(k^{k^6})$ new ones, that respond to the guessing in the second step, etc.
After creating all of the guesses we end up with exponential in $k$ different cases.
We now create one partial ILP instance for each case.
From now on we assume that we know exactly the structure of all paths we have guessed and the permutation of all variables.

\subsection{Properties of 
\texorpdfstring{\deltaExact } {Simple Delta-TGR}
}

In this section we study the properties of our problem, that then help us creating constraints of our ILP.
In our analysis we exploit the following greatly, that is why we state is as an observation.

\begin{observation}\label{obs:FirstLabelAndDuration}
    Let $P$ be the underlying path of a fastest temporal path from $u$ to $v$, where $e_1, e_p \in P$ are its first and last edge, respectively.
    Then, knowing the label $\lambda (e_1)$ of the first edge and the duration $d(P,\lambda)$ of the temporal path $(P,\lambda)$, we can uniquely determine the label $\lambda (e_p)$ of the last edge of $P$.
    Symmetrically, knowing $\lambda(e_p)$ and $d(P)$, we can uniquely determine $\lambda(e_1)$.
\end{observation}
The correctness of the above statement follows directly from \cref{def:temporalPath+Duration}. Since the duration of $(P,\lambda)$ is calculated as the difference of labels of last and first edge plus $1$,
where the label of last edge is considered with some delta periods,
\ie $d(P,\lambda) = p_i \Delta + \lambda(e_p) - \lambda (e_1) + 1$, for some $p_i \geq 0$.
Therefore $d(P,\lambda) \pmod \Delta \equiv  (\lambda(e_p) - \lambda (e_1) + 1) \pmod \Delta$.

@@

\begin{claim}\label{claim:FPT-uv-Labelalledges}
    Let $u, v \in U$ be two arbitrary vertices of interest and suppose that $P = (u=v_1,v_2, \dots, v_p = v)$, where $p \geq 2$, 
    is a $u,v$ path in $G'$, which is also the underlying path of a fastest temporal path from $u$ to $v$.
    Moreover suppose also that $P$ is a segment.
    %Suppose also that no other vertex from $P$ is a vertex of interest, \ie $P \setminus \{u,v\} \cap U = \emptyset$.
    We can determine the labeling $\lambda$ of every edge in $P$ with respect to the label $\lambda(uv_2)$ of the first edge.
    %such that the labeling $\lambda$ respects $D'$.
\end{claim}

\begin{proof}
We claim that $u$ reaches all of the vertices in $P$ the fastest, when traveling along $P$ (\ie by using a subpath of $P$).
To prove this suppose for the contradiction that there is a vertex $v_i \in P \setminus \{u,v\}$, that is reached from $v$ on a path different than $P_i = (u, v_2, v_3, \dots, v_i)$.
Since the only vertices of interest of $P$ are $u$ and $v$, it follows that all other vertices on $P$ are of degree $2$. 
Then the only way to reach $v_i$ from $u$, that differs from $P$, would be to go from $u$ to $v$ using a different path $P_2$,
and then go from $v$ to $v_{p-1}, v_{p-2}, \dots, v_i$.
But since $P$ is the fastest temporal path from $u$ to $v$, we get that $d(P_2) \geq d(P)$ and $d(P_2 + (v,v_{p-1}, \dots, v_i) > d(P) > d(P_i)$.

@@

%Observe also the following.

\begin{claim}\label{claim:FPT-unlabeldPaths-01}
    Suppose that
    $S_{u,v}, S_{w,z}$ are two segments with $ v_i \in S_{u,v}$ and $z_j \in S_{w,z}$,
    where $z_j$ is a split vertex of $v_i$ in the segment $S_{w,z}$.
    Suppose that the fastest temporal path from $v_i$ to $z_j$ travels through vertices $u$ and $w$.
    Then the fastest temporal path from $v_i$ to any other vertex of $S_{w,z}$, that is closer to $w$,
    travels through the same two vertices $u$ and $w$.
    Similarly it holds for the cases when the fastest temporal path travels through $w,v$ or $z,u$ or $z,v$.
\end{claim}
\begin{proof}
    Let $z_\ell$ be a vertex of $S_{w,z}$, that is closer to $w$ than $z$ in the segment.
    Let us denote with $P_{v_i,z_j}$ the underlying path of the fastest temporal path from $v_i$ to $z_j$.
    Denote with $P_{v_i,z_j}^\ell$
    the subpath of the fastest temporal path from $v_i$ to $z_j$, that terminates in $z_\ell$.
    We want to show that $P_{v_i,z_j}^\ell$ is an underlying path of a fastest temporal path from $v_i$ to $z_j$.
    Let us observe the following possibilities.
    \\
    First, suppose for the contradiction, that the fastest temporal path from $v_i$ to $z_\ell$ travels through vertices $u$ and $z$.
    Denote this path as $P^1_{v_i,z_\ell}$.
    Then it follows that $d(P^1_{v_i,z_\ell}, \lambda) \leq d(P_{v_i,z_j}^\ell,\lambda)$, which would imply that
    the duration of the temporal path from $v_i$ to $z_j$ using the subpath of $P^1_{v_i,z_\ell}$, would be strictly smaller than the duration of $(P_{v_i,z_j},\lambda)$, which cannot be possible.
    \\
    Second, suppose that the fastest temporal path from $v_i$ to $z_\ell$ travels through vertices $v$ and $w$.
    Denote this path as $P^2_{v_i,z_\ell}$.
    Note that $P_{v_i,z_j}^\ell$ and $P^2_{v_i,z_\ell}$ intersect on a segment $S_{w,z}$ from the vertex $w$ to $z_ell$.
    Therefore since
    $d(P^2_{v_i,z_\ell}, \lambda) \leq d(P_{v_i,z_j}^\ell,\lambda)$,
    and since there is unique way to
    extend the path $P^2_{v_i,z_\ell}$ from $z_\ell$ to $z_j$, denote the extended path as $P^j_{v_i,z_\ell}$,
    we get that $d(P^j_{v_i,z_\ell}, \lambda) \leq $$(P_{v_i,z_j},\lambda)$.
    Which implies that $d(P^j_{v_i,z_\ell}, \lambda) = d(P_{v_i,z_j}, \lambda)$.
    Now using the similar argument it follows that $d(P_{v_i,z_j}^\ell, \lambda) = d(P^2_{v_i,z_\ell}, \lambda)$,
    therefore $P_{v_i,z_j}^\ell$ is also a fastest temporal path from $v_i$ to $z_j$.
    \\
    Third, suppose that the fastest temporal path from $v_i$ to $z_\ell$ travels through vertices $v$ and $z$.
    Denote this path as $P^3_{v_i,z_\ell}$.
    Then the duration of the temporal path from $v_i$ to $z_j$ using the subpath of $P^3_{v_i,z_\ell}$, would be strictly smaller than the duration of $(P_{v_i,z_j},\lambda)$, which cannot be possible.
\end{proof}

\subsection{Adding constraints to the ILP}
We proceed with adding constraints to our ILP instance, by iterating through all pairs of vertices.
We now describe how we add constraints to a path. Whenever we say that a duration of a path gives an equality or inequality constraint, we mean the following.
Let $P=(u=v_1,v_2, \dots, v_p = v)$ be the underlying path of a fastest temporal path from $u$ to $v$,
and let $Q = (u=z_1,z_2, \dots, z_r = v)$ be the underlying path of another temporal path from $u$ to $v$.
Then we know that $d(P,\lambda) = D_{u,v}$ and $d(Q, \lambda) \geq D_{u,v}$.
Using \cref{obs:durationPwithWaitingTimes}
we create an \emph{equality constraint} for $P$
of the form 
\begin{equation}\label{eq:FPT-equalityConstraint}
    D_{u,v} = \sum _ {i=2} ^ {p-1} (\lambda (v_{i}v_{i+1}) - \lambda (v_{i-1}v_i))_\Delta + 1,
\end{equation}
and an \emph{inequality constraint} for $Q$ 
\begin{equation}\label{eq:FPT-inequalityConstraint}
    D_{u,v} \leq \sum _ {i=2} ^ {r-1} (\lambda (z_{i}z_{i+1}) - \lambda (z_{i-1}z_i))_\Delta + 1.
\end{equation}
In both cases we implicitly assume that if the difference is smaller than zero we add the value $\Delta$ to it (\ie we consider the difference modulo $\Delta$), therefore we have the sign $\Delta$ around the brackets.

First case we observe is the case when we want to determine the labels on fastest paths from $u$ to $v$, where $u,v \in U$.

@@

We already studied the case of fastest paths starting and ending in a vertex of $U$,
so now we have to study all other cases.
